---
title: "Homework Assignment 1"
author: "Zejie Gao"
date: "2023-02-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The dataset trees contains measurements of Girth (tree diameter) in inches, Height in feet, and Volume of timber (in cubic feet) of a sample of 31 felled black cherry trees. The following commands can be used to read the data into R.
```{r}
require(datasets)
head(trees)
```

(a) Briefly describe the data set trees, i.e., how many observations (rows) and how many variables (columns) are there in the data set? What are the variable names?
```{r}
nrow(trees)
ncol(trees)
ls(trees) 
# There are 31 rows and 3 column in the data set, with three variables named "Girth","Height", and  "Volume"
```
(b) Use the pairs function to construct a scatter plot matrix of the logarithms of Girth, Height and Volume.
```{r}
library(tidyverse)
log_trees <- trees |>
  mutate(Girth = log(Girth)) |>
  mutate(Height = log(Height))|>
  mutate(Volume = log(Volume))
pairs(log_trees)
```
(c) Use the cor function to determine the correlation matrix for the three (logged) variables.
```{r}
cor(log_trees)
```
(d) Are there missing values?
```{r}
is.na(log_trees)
sum(is.na(log_trees))
# No, there is no missing values
```
(e) Use the lm function in R to fit the multiple regression model:
    log(Volumei) = beta0 + beta1 log(Girthi) + beta2 log(Heighti) + ei
    and print out the summary of the model fit.
```{r}
fit <- lm(Volume ~ Girth + Height, data = log_trees)
fit
y <- log_trees$Volume
x1 <- log_trees$Girth
x2 <- log_trees$Height
R.2 <- 1 - sum((fit$residuals^2))/ (sum((y - mean(y))^2))
R.2
summary(fit)
# Estimation of the lm: log(Volumei) = -6.63 + 1.98 log(Girthi) + 1.12 log(Heighti) + ei
# Since R^2 (0.9777 or 0.978 on the summary) is very close to 1, the model better fits the data. 
```
(f)

```{r}
v1 <- log_trees$Girth
v2 <- log_trees$Height
X <- cbind(rep(1, times=nrow(log_trees)),v1,v2) 
y <- matrix(log_trees$Volume)
beta_hat <- solve(t(X)%*%(X))%*%(t(X)%*%(y))
beta_hat
# The beta_hat matrix match the output I got in (e)
```
(g) Compute the predicted response values from the fitted regression model, the residuals, and an estimate of the error variance.

```{r}
y_hat <- X%*%beta_hat
y_hat
Res <- fit$residuals
sigma2.hat <- sum(Res^2) / fit$df.residual
sigma2.hat
SSR <- sum(fit$residuals^2)
SSR
# SSR = 0.1855
# Var(ei) = 0.006624
```
2. Consider the simple linear regression model:
                                
(a) Assume beta0 = 0. What is the interpretation of this assumption? What is the implication on the regression line? What does the regression line plot look like?                                
```{r}
# The assumption represent that the yi will have a high likelihood to be zero when xi equals to zero. 
# It implicate the y intercept of linear regression model is zero. 
# The regression line start from coordinate (0,0).
```
(b) Derive the LS estimate of beta1 when beta0 = 0.
```{r}
# beta1 = sum((y-mean(y))*(x-mean(x)))/sum((x-mean(x))ˆ2)
# beta1 does not influenced by beta0 value
```
(c) How can we introduce this assumption within the lm function?
```{r}
# lm(y ~ x-1, data = dataset)
# Based on the assumption beta0 = 0, beta0 from lm function can be deleted.Then, the lm function will become yi = beta1xi +ei 
```
(d) For the same model, assume beta1 = 0. What is the interpretation of this assumption? What is the implication on the regression line? What does the regression line plot look like?
```{r}
# The assumption represent that the value of yi does not affected by the value of xi. The predictor x1 is not significant when explaining yi.
# It implicate the linear regression model wasn't able to find a linear relationship between the yi and x1.
# The plot will only have a horizontal line which is y = constant value (beta0 + ei) or so called the mean of y.
```
(e) Derive the LS estimate of beta0 when beta1 = 0.
```{r}
# beta0 = mean(y)
```
(f) How can we introduce this assumption within the lm function?
```{r}
# lm(y ~ 1, data = dataset)
# Based on the assumption beta1 = 0, beta1 part from lm function can be deleted.Then, the lm function will become  y1i = beta0 +  ei.
```
3. Consider the simple linear regression model:
                             
(a) Use the LS estimation general result beta_hat =  to find the explicit estimates for beta0 and beta1.
```{r}
# on the pdf
```

(b) Show that the LS estimates beta0_hat and beta1_hat are unbiased estimates for beta0 and beta1 respectively.
```{r}
# on the pdf
```









